Pre-/Post-Trikafta (Prospective)

### 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

All metrics calculated based on 2 hour

```{r}
# convert Insulin to mcU/mL
ins_cols = colnames(df)[grep("insulin",tolower(colnames(df)))]
df[,ins_cols] = lapply(df[,ins_cols], function(c){
  return(as.numeric(c) * 0.02869)
})
# cpeptide to pmol/L
cpep_cols = colnames(df)[grep("c.peptide",tolower(colnames(df)))]
df[,cpep_cols] = lapply(df[,cpep_cols], function(c){
  c = as.numeric(c)/1000
  return(c * 0.33)
})
# Calculate AUCs
df$`Glucose AUC` = apply(df %>% select(Glucose.0.minute:Glucose.120.min),1,function(r){
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
df$`C Peptide AUC` = apply(df %>% select(c.peptide.0.minute:c.peptide.120.min),1,function(r){
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
# Other indices
# Cpeptide index = delta cpeptide 0-30/delta glucose 0-30
df$`C Peptide Index` = (df$c.peptide.30.min - df$c.peptide.0.minute)/
  (df$Glucose.30.min - df$Glucose.0.minute)
# Add labels
newlabels["A1C.result"] = "HbA1c (%)"
p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+
              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`, 
            data = df, id = Study.ID)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

>>>>

All metrics calculated based on 2 hour OGTT (0, 10, 30, 60, 90, and 120 minutes).

```{r results='asis'}
# Make dataset for HOMA2 calculations - glucose in mmol/L and cpep in nmol/L
homa = df %>% select(Study.ID,Event.Name,Glucose.0.minute:Glucose.180.min,
                     c.peptide.0.minute:c.peptide.180.min)
glu_cols = colnames(df)[grep("glucose\\.\\d{1,}",tolower(colnames(df)))]
cpep_cols = colnames(df)[grep("c.peptide",tolower(colnames(df)))]
homa[,glu_cols] = lapply(homa[,glu_cols], function(c){
  c/18
})
homa[,cpep_cols] = lapply(homa[,cpep_cols], function(c){
  c = suppressWarnings(as.numeric(c))/1000
  c*0.33
})
write.csv(homa,"./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/homa_data.csv",
          na = "",row.names = F)
# Glucose tolerance group
glu_cols = c("Glucose.0.minute","Glucose.60.min","Glucose.120.min")
df$Group = apply(df[,glu_cols],1,function(r){
  if(r[1]<126 & r[2] < 200 & r[3] < 140){
    return("NGT")
  } else if (r[1] >= 126 | r[3] >= 200){
    return("CFRD")
  } else {
    return("AGT")
  }
})
# convert Insulin to mcU/mL
ins_cols = colnames(df)[head(grep("insulin",tolower(colnames(df))),-2)]
df[,ins_cols] = lapply(df[,ins_cols], function(c){
  return(as.numeric(c) * 0.02869)
})
# cpeptide to pmol/L
cpep_cols = colnames(df)[head(grep("c.peptide",tolower(colnames(df))),-2)]
df[,cpep_cols] = lapply(df[,cpep_cols], function(c){
  return(c * 0.33)
})
# Calculate AUCs
df$`Glucose AUC` = apply(df %>% select(Glucose.0.minute:Glucose.120.min),1,function(r){
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
df$`C Peptide AUC` = apply(df %>% select(c.peptide.0.minute:c.peptide.120.min),1,function(r){
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
# Incremental AUCs
df$`Glucose iAUC` = apply(df %>% select(Glucose.0.minute:Glucose.120.min),1,function(r){
  r = r - r[1]
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
df$`C Peptide iAUC` = apply(df %>% select(c.peptide.0.minute:c.peptide.120.min),1,function(r){
  r = r - r[1]
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
# Other indices
# Cpeptide index = delta cpeptide 0-30/delta glucose 0-30
df$`C Peptide Index` = (df$c.peptide.30.min - df$c.peptide.0.minute)/
  (df$Glucose.30.min - df$Glucose.0.minute)
# Matsuda c peptide (500,000/√[fasting glucose*fasting cpep*mean glucose*mean cpep])   [Glucose mg/dL, cpeptide pmol/L]
glu_cols = colnames(df)[head(grep("glucose\\.\\d{1,}",tolower(colnames(df))),-2)]
df$mean_glucose = rowMeans(df[,glu_cols])
df$mean_cpep = rowMeans(df[,cpep_cols])
df$`Adapted Matsuda` = 500000/(sqrt(df$Glucose.0.minute*df$c.peptide.0.minute*
                                      df$mean_glucose*df$mean_cpep))
# C Pep ODI
df$`C Peptide ODI` = df$`C Peptide Index`/df$`Adapted Matsuda`
# iAUC ratio
df$iAUC_ratio = df$`C Peptide iAUC`/df$`Glucose iAUC`
# Add labels
newlabels["A1C.result"] = "HbA1c (%)"
newlabels["Glucose.0.minute"] = "Fasting glucose (mg/dL)"
newlabels["Glucose.60.min"] = "1hr glucose (mg/dL)"
newlabels["Glucose.120.min"] = "2hr glucose (mg/dL)"
newlabels["c.peptide.0.minute"] = "Fasting C-peptide (pmol/L)"
newlabels["iAUC_ratio"] = "C pep iAUC:Glucose iAUC"
p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+Group+Glucose.0.minute+
              Glucose.60.min+Glucose.120.min+c.peptide.0.minute+
              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`+`Adapted Matsuda`+
              `C Peptide ODI`+iAUC_ratio, 
            data = df, id = Study.ID)
summary(p,pfootnote = T,labelTranslations = newlabels)
```
